---
title: "ANZ EDA"
output: html_notebook
---
## Load data

```{r}
data <- readxl::read_xlsx("F:/ANZ synthesised transaction dataset.xlsx")
head(data)
```
## Column names
```{r}
names(data)
```
## Diagnose dataset
```{r}
library(dlookr)
diagnose(data)

```
Convert variables that should be factors to factors
```{r}
data$status <- as.factor(data$status)
data$card_present_flag <- as.factor(data$card_present_flag)
data$country <- as.factor(data$country)
```

## Exploratory Data Analysis
### Spending by location
```{r}
library(ggplot2)
library(dplyr)
library(tidyr)

x <- data[data$movement==c("debit"),] %>%
  select(transaction_id, long_lat) %>%
  mutate(s = strsplit(long_lat, " ")) %>%
  mutate(s = purrr::map(s, setNames, c("Longitude", "Latitude"))) %>%
  unnest_wider(s)
  
library(rworldmap)
newmap <- getMap(resolution = "high")
# plot map
plot(newmap, xlim = c(112,155), ylim = c(-40, -10), 
     border = "darkgray", 
     col = "wheat2", bg = "gray95")
points(x$Longitude, x$Latitude, col = alpha("red",.01), pch =20, cex = 1 )
    
```
Transactions are mostly concentrated to the Eastern coast and western coast. Expanding along the coast should be looked into as well as increasing marketing in these regions.

### Spending by age
```{r}
Total_Traxcn <- aggregate(data[data$movement==c("debit"),]$amount, by = list(Age=data[data$movement==c("debit"),]$age), FUN = sum)
ggplot(Total_Traxcn, aes(y = x, x =Age)) + geom_col() + ylab("Total Amounts Transacted by Age")
```
```{r}
sum(Total_Traxcn[Total_Traxcn$Age<=30,]$x)/sum(Total_Traxcn$x)
```
```{r}
sum(Total_Traxcn[(Total_Traxcn$Age>30 & Total_Traxcn$Age<50),]$x)/sum(Total_Traxcn$x)

```
53 percent of transactions are done by people under the age of 30 whilst 43 percent are people between 30 and 50.
Thus, this demographic is a key domegraphic and efforts should be made towards engaging users to expand users belonging to this demographic.


### Account Balance by Age 
```{r}
ggplot(data, aes(y = balance, x  = age)) + geom_boxplot(aes(group = age))
```
Average balance increases with age. Middle aged demographics can be targeted for investing these funds into mutual funds.

### Payments by Day
```{r}
library(lubridate)
data$weekday <- wday(data$extraction)
data$monthday <- mday(data$extraction)
weekday <- data[data$movement==c("debit"),] %>% 
  select(amount)%>%
  aggregate(by=list(data[data$movement==c("debit"),]$weekday), FUN = mean)
monthday <- data[data$movement==c("debit"),] %>% 
  select(amount)%>%
  aggregate(by=list(data[data$movement==c("debit"),]$monthday), FUN = mean)
par(mfrow=c(1,2))
plot(x = weekday$Group.1, y = weekday$amount, xlab= c("Day of The Week"), ylab = c("Average Amount Spent"), 'l')

plot(x = monthday$Group.1, y = monthday$amount, xlab= c("Day of The Month"), ylab = c("Average Amount Spent"), 'l')

```
People spend the least on Saturdays and expenditure increases towards the middle of the month.



